---
title: "Regression Models Course Project"
author: "Han Yunxi"
date: "18 February 2015"
output: pdf_document
---
## Executive Summary

*Motor Trend*, a magazine about the automobile industry, are interested in exploring the relationship between a set of automobile variables and miles per gallon *(mpg)*. They are particularly interested in the following two questions:

<li>Is an automatic or manual transmission better for MPG?
<li>Quantify the MPG difference between automatic and manual transmissions.

The statistical inference shows that from the results, we see that manual cars average 24.4 *mpg*, automatic cars average 17.1 *mpg*. We also fit 4 multiple regression models, to find the model with the highest $R^2$ value. The final regression equation (model 4) is **mpg = 9.72 + (-2.93)wt + 1.01qsec + 14.08am + (-4.14)wt:am**. The conclusion is that cars that are lighter in weight with a manual transmission will have higher MPG values.

## Exploratory Data Analysis (EDA)
From the `?mtcars` help file, we see that the dataset comprises fuel consumption *(outcome)* and 10 aspects *(predictors)* of automobile design and performance for 32 automobiles (1973–74 models). 

```{r warning=FALSE}
library(data.table)
library(knitr)
data(mtcars)
```

#### EDA Findings
A scatterplot matrix of *mpg* and other 10 predictors is generated. The top row shows the relationship between *mpg* and each of the predictors. From the scatterplot, there appears to be some relationship  between *mpg* and *disp* , *hp*, *wt*. There are a great deal of random variables present in other predictors. In other words, the model is likely to consist of *disp* , *hp* and *wt*. This requires further investigation. *Refer to appendix for scatterplot*

## Statistical Inference
Here, we calculate the 95% confidence intervals on *mpg* for automatic and manual cars. A t-test is used to evaluate the if there is no difference in average *mpg* between automatic($\mu_0$) and manual($\mu_a$) cars at 95% confidence interval. $H_0$ : $\mu_0$ = $\mu_a$ vs. $H_a$ : $\mu_0$ < $\mu_a$ 

```{r}
  t.test(mpg ~ am, data = mtcars, alternative = c("less"),paired=FALSE, conf.level = 0.95) 
```
From the results, we see that manual cars average 24.4 *mpg*, automatic cars average 17.1 *mpg*. Since the p-value, 0.0006 is lesser than 0.05, we reject $H_0$ and conclude that manual cars have a better average  *mpg* than automatic cars.

## Multiple Regression
We will start with the predictors given in EDA (model 1), then with full model (model 2) that predicts *mpg* based on all predictors given in mtcars dataset. Thereafter we use **step** function to eliminate predictors with the highest p-value and refit the model to arrive at the final model (model 3). *All results can be found in appendix*

###Model 1
```{r results='hide'} 
fit_EDA <- lm(mpg ~ disp + hp + wt, data = mtcars)  #model 1
```
Using the EDA results, we fit **mpg = disp + hp + wt** and test if it is a good fit. The adjusted $R^2$ is 0.8. So about 80% of the variation in the scores can be predicted using the model. Let's see if the model can be improved further.

###Model 2
```{r results='hide'}
fit_Full <- lm(mpg ~ ., data=mtcars)  # model 2
```
The results for model 2 shows the adjusted $R^2$ is also 0.8. Similar to Model 1, about 80% of the variation in the scores can be predicted using the model. There is no improvement to fitting all predictors.
 
###Model 3
```{r results='hide'}
fit_Stepwise <- step(lm(mpg ~ ., data=mtcars)) # model 3
```
The results shows that adjusted $R^2$ value 0.83 - an improvement of 3% over the full model. P-values are also significant at 95%. This model **mpg = wt + qsec + am** seems to be a good fit. 

###Interactions (model 4)
To overcome the limitation of **step** function to account for Interactions, we will introduce (wt*am) term to the regression model (model 3). It would be useful to add an interaction term to the model if we wanted to test the hypothesis that the relationship between the weight(wt) on the mpg was different for manual rather than automatic cars.

```{r}
fit_Interaction <- lm(mpg ~ wt + qsec + am + wt*am , data = mtcars) 
```

The adjusted $R^2$ for model 4 is 0.88, which is the best by far. Hence the final regression equation (model 4) will be **mpg = 9.72 + (-2.93)wt + 1.01qsec + 14.08am + (-4.14)wt:am**. 

###Residual Diagnostics
After selecting regression values and fitting the regression model 4, it is necessary to plot the residuals to investigate good/poor model fit. The results are 1. The Residuals vs. Fitted plot shows no systematic patterns and the variation in the residuals does not seem to change with the size of the fitted values. 2. The histogram of residuals is normally distrubuted.

##Results
To interpret the results, 
<li>1. For every 1000lb increase in weight, the *mpg* for manual transmission shows a $14.08(1) - 4.14(wt)(1) = 9.94$ improvement over automatics.
2. Manual transmission is better for *mpg*.

-- end --


## Appendix

###Exploratory Data Analysis
```{r echo=FALSE}
pairs(mtcars, panel=panel.smooth,main="Pair Graph of Motor Trend Car Road Tests")
```

###Results of Multiple Regression
```{r}
summary(fit_EDA)$coef
summary(fit_Full)$coef
summary(fit_Stepwise)$coef
summary(fit_Interaction)$coef

#calculating R-squared values
a<-summary(fit_EDA)$adj.r.squared
b<-summary(fit_Full)$adj.r.squared
c<-summary(fit_Stepwise)$adj.r.squared
d<-summary(fit_Interaction)$adj.r.squared
R <- c(a,b,c,d)
R
```

###Residual Diagnostics
```{r echo=FALSE}
par(mfrow=c(2,2))
plot(mtcars$wt, residuals(fit_Interaction), xlab="wt")
plot(mtcars$qsec, residuals(fit_Interaction), xlab="qsec")

plot(fitted(fit_Interaction), residuals(fit_Interaction), 
     xlab = "Predicted mpg", ylab ="Residuals", main = "Residuals vs Fitted")

hist(residuals(fit_Interaction), main="Histogram of Residuals")
```
### http://hyunxi.github.io/JHU-DataScience/Module%207-Regression%20Models/Course_Project.html
